---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Archana Vijayakumar Sreekala"
date: "September, 27 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false

install.packages("tidyverse")
install.packages("tidymodels")
install.packages("gt")
install.packages("gtExtras")
install.packages("DataExplorer")

library(magrittr)     # the pipe
library(tidyverse)    # for data wrangling + visualization
library(tidymodels)   # for modeling
library(gt)           # for making display tables
library(gtExtras)     # helper functions for beautiful tables
library(DataExplorer) #

the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

## Exercises

### Exercise 1

The `the_tate` dataset has \_\_\_ unique artists who worked from \_\_\_ to \_\_\_. The works were acquired between the years \_\_\_ and \_\_\_.

```{r Data Analysis}

the_tate
the_tate_artists

dim(the_tate)
dim(the_tate_artists)

the_tate |> dplyr::slice_head(n=2) |> dplyr::glimpse()
the_tate_artists |> dplyr::slice_head(n=2) |> dplyr::glimpse() 

view(the_tate)
view(the_tate_artists)

summary(the_tate)

# Summarise function to group by each artist and find the summarised information 
the_tate |> dplyr::group_by(artist) |>
  dplyr::summarise(N=n(), 
         min_year = min(year, na.rm = TRUE),          
         max_year = max(year, na.rm = TRUE), 
         min_acquisition_year = min(acquisitionYear, na.rm = TRUE),
         max_acquisition_year = max(acquisitionYear, na.rm = TRUE)
  )

library(dplyr)
# To find the number of unique artists

unique_artist_count <- 
  the_tate |> select(artist) |>
         distinct() |>
         nrow() 
# To find the range of years that the artists worked
min_year <- min(the_tate$year, na.rm = TRUE)         
max_year <- max(the_tate$year, na.rm = TRUE)  
# To find the range of acquisition years
min_acquisition_year <- min(the_tate$acquisitionYear, na.rm = TRUE)
max_acquisition_year <- max(the_tate$acquisitionYear, na.rm = TRUE)

cat("The the_tate dataset has", unique_artists_count, "unique artists who worked from",
    min_year, "to", max_year, ".\n The works were acquired between the years", min_acquisition_year, "and", max_acquisition_year, ".\n")


DataExplorer::introduce(the_tate)
DataExplorer::introduce(the_tate_artists)

DataExplorer::plot_missing(the_tate) 
DataExplorer::plot_missing(the_tate_artists)
```

### Exercise 2

How number of works with missing dates is \_\_.

The number of artists whose works have missing dates is \_\_.

It would require resolving missing year data for only \_\_ artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as \_\_\_\_.

### Exercise 3

The artist with the most works in the Tate collection is \_\_\_.

The artist with the tenth-most works in the Tate collection is \_\_\_.

### Exercise 4

The artist with the greatest number of works in the Tate collection represent \_\_\_% of the total number of works

### Exercise 5

There are \_\_ duplicate artist-title pairs

### Exercise 6

The artist with the largest work in the tate collection is \_\_\_

The artist with the smallest work in the collection is \_\_\_. The smallest work has area \_\_\_ $\text{cm}^2$

### Exercise 7

...

### Exercise 8

The annual return in the SPX price in 2020 was \_\_\_%.

The corresponding price volatility was \_\_\_%.

### Exercise 9

The period volatility was \_\_\_.\_%

### 
